A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-05-18 meth.tex
using all models for cov figure

@@

While historically expert opinion has guided the determination of harmonization
methods to choose for a given trajectory, a new, \textit{decision tree} approach
has been implemented in this work. In order to provide reasonable
\textit{default} methods, the historical trajectory, unharmonized model
trajectory, and relative difference between history and model values in the
harmonization year are analyzed. A figure of the decision tree used in this
analysis is shown in Figure \ref{}.

% decision tree figure

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is possible for models to report zero values in the
harmonization year in situations in which technologies are introduced in future
time periods in regions or for sectors which produce an emissions species that
is absent in the initial modeling period. In such cases, an offset method is
required as a ratio method would mask future emissions and erroneously harmonize
the trajectory.

Variability of the historical trajectory is also an important characteristic
when considering the choice of harmonization method. Such trajectories are
normally indicative of emissions from land use sectors. Land use sectors are
modeled with varying fidelity by IAMs, but in general proxies are used for
emissions trajectories. Take for example the emissions of black carbon (BC) from
grassland burning in Africa \TODO{check that this is a good sector}: for most
models, the use of grass and paturelands is modeled, and the emissions of
species from the burning of these lands is determined by applying a emissions
factor to a proxy model of burning of these lands; in general, the more
grassland usage occurs, the higher the emissions from burning of
grasslands. Therefore, consistency in harmonization method is important due to
the approaches taken in modeling these sectors by IAMs. In order to reliably
detect land-use-like emissions, a measure of the covariance of the historical
trajectory is calculated as shown in Equation \ref{eqs:cov} and tested against a
threshold. To determine this threshold, an analysis of the recent CEDS
\TODO{cite ceds paper} historical data has been performed. Figure \ref{fig:cov}
shows the distribution of land-use covariances and non-land-use covariances as
determined for \TODO{list regions for 11 regions}. The minimum covariance
observed for land-use sectors is \TODO{find}, thus theshold used in this study
is \TODO{do this}. Notably, some non-land-use sectors , however,

@@

>>>>

\subsection{Harmonization Methods}

@@

\subsection{Default Method Decision Tree}

While historically expert opinion has guided the determination of harmonization
methods to choose for a given trajectory, a new, \textit{decision tree} approach
has been implemented in this work. In order to provide reasonable
\textit{default} methods, the historical trajectory, unharmonized model
trajectory, and relative difference between history and model values in the
harmonization year are analyzed. A figure of the decision tree used in this
analysis is shown in Figure \ref{fig:decision_tree}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{decision_tree.png}
    \caption[]{
      \label{fig:decision_tree}
      The default method decision tree used in the \code{aneris} software
      library. Decision-making threshold are described below, and the methods
      applied are described in Section \ref{sec:impl-meths}.}
  \end{center}
\end{figure}

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is possible for models to report zero values in the
harmonization year in situations in which technologies are introduced in future
time periods in regions or for sectors which produce an emissions species that
is absent in the initial modeling period. In such cases, an offset method is
required as a ratio method would mask future emissions and erroneously harmonize
the trajectory. In most cases, however, models do report values in the
harmonization year. When model and historical values are relatively close, a
convergence method is provided in order to be as representative as possible to
the underlying unharmonized model results. The final model-based characteristic
of import is whether the model reports negative emissions. Such a case is
possible for gas species which can be extracted from the enviornment, as is the
case for CO2 in future scenarios with high mitigation strategies.

Variability of the historical trajectory is also an important characteristic
when considering the choice of harmonization method. Such trajectories are
normally indicative of emissions from land use sectors. Land use sectors are
modeled with varying fidelity by IAMs, but in general proxies are used for
emissions trajectories. Take for example the emissions of black carbon (BC) from
grassland burning in Africa \TODO{check that this is a good sector}: for most
models, the use of grass and paturelands is modeled, and the emissions of
species from the burning of these lands is determined by applying a emissions
factor to a proxy model of burning of these lands; in general, the more
grassland usage occurs, the higher the emissions from burning of
grasslands. Therefore, consistency in harmonization method is important due to
the approaches taken in modeling these sectors by IAMs. In order to reliably
detect land-use-like emissions, a measure of the covariance of the historical
trajectory is calculated as shown in Equation \ref{eqs:cov} and tested against a
threshold. To determine this threshold, an analysis of the recent CEDS
\TODO{cite ceds paper} historical data has been performed. Figure \ref{fig:cov}
shows the distribution of land-use covariances and non-land-use covariances as
determined for historical data aggregated to 11 world regions \TODO{list regions
  for 11 regions}. A threshold value of 10 has been chosen based on the
11-region observations, but, importantly, any regional aggregation is possible
as this is model depedent.

@@

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization year.

\subsection{Implemented Methods}\label{sec:impl-meths}
